A demonstration of the \LaTeXe\ class file for \itshape{SAGE Publications}

### 2015-11-30 body/intro.tex
intro + overview no fig

The issue of planning acyclic contacts was first completely described by \citeauthor{conf/iser/BretlRLKA04}, where  it is proven to require the handling of two simultaneous problems, $\mathcal{P}_1$: a relevant guide trajectory for the root of the robot in $SE(3)$; and $\mathcal{P}_2$: the planning of a discrete sequence of acyclic, balanced contact configurations along the trajectory\endnote{A third non trivial problem, $\mathcal{P}_3$, not adressed in this work, then consists in interpolating a complete motion between two postures of the contact sequence. Noticebly, \citeauthor{Carpentier2016} combine
our planner with such a method to generate the complete motion.}.  A key issue is to avoid combinatorial explosion when considering at the same time the possible contact configurations and the potential trajectories. This seminal paper proposes a first effective algorithm, able to handle simple situations (such as climbing scenarios), but not scalable to arbitrary environments. Following it, seve\-ral papers have applied this approach in particular situations, typically limiting the combinatorial by imposing a fixed set of possible contacts \citep{Hauser06usingmotion, stilman2010}.

Most of the papers that followed the work of \citeauthor{conf/iser/BretlRLKA04} have explored alternative formulations to handle the combinatorial issue. Two main directions have been explored. \textbf{On one hand, local optimization of both the root trajectory \Pa and the contact positions $\mathcal{P}_2$} has been used, to trade the combinatorial of the complete problem for a differential complexity, at the cost of local convergence. A complete example of the potential offered by such approaches was propose \citep{Mordatch:2012:DCB:2185520.2185539} and successfully applied to a real robot \citep{mordatch2015}. To keep reasonable computation times, the method uses a simplified dynamic model for the avatar. Still, the computation time is far from interactive  (about 1 minute of computation for a sequence of 20 contacts). \citeauthor{DBLP:conf/humanoids/DeitsT14} propose to solve contact planning globally as a mixed integer problem, but only cyclic, bipedal locomotion is considered. Aside from the computation cost, a major drawback of these optimization based approaches is thus that they only offer local convergence when applied to acyclic contact planning.

@@

\subsection{Paper contribution and organization}
Although the optimization approach is promising, we choose to focus on the sample-based methodology, more able to find complex trajectories in cluttered environments. While the theoretical structure of the problem is well understood, there is currently no scalable method to solve it. The combinatorial of the original problem (as described by \citeauthor{conf/iser/BretlRLKA04}) is too high to have any hope of tackling it directly. Alternative formulations are necessary to obtain practical solutions. We believe that the separation between the guide trajectory and the contact sequence is the most promising direction \citep{DBLP:conf/iser/EscandeKMG08}. However, this direction raises two theoretical questions that remain to be solved, or even to be properly formulated:
\begin{itemize}
\item The guide trajectory must satisfy a property guaranteeing the existence of a contact sequence to actuate it\endnote{This property is related to the controllability of the root actuated by the contact forces, but for discrete bounded properties.}. This property has not been studied yet: the only way to validate a trajectory is to explicitly compute the contacts, which is computationally not reasonable \citep{Bouyarmane2009}.
\item There is an infinite combination of possible contact sequences for a given root trajectory. The selection of one particular contact sequence with interesting properties (minimum number of contact change, robustness, efficiency or naturalness) has been studied for cyclic cases \citep{Hauser06usingmotion}, but has not been efficiently applied to cluttered environments (\citeauthor{bouyarmane:lirmm-00777727, DBLP:conf/iser/EscandeKMG08} mostly randomly pick one contact sequence, leading to possibly very tedious contact sequences).

\end{itemize}
\begin{figure*}[t]
  \centering
  \begin{overpic}[width=1\linewidth]{figures/workflow_new}
    \put (6.4,1.8) {\normalsize{Request}} 
    \put (22,11.3) {\large{A} }
    \put (31.6,2) {\scriptsize{\color{white}RB-PRM}} 
    \put (46.4,11.3) {\large{B} }
    \put (72.5,2) {\tiny{\color{white}EFORT}} 
  \end{overpic}
  \vspace{-1em}
  \caption{
    Overview of our 2-stage framework. (A) Given a path request between the yellow and blue positions, a guide trajectory is computed in $C_{reach}$ using RB-PRM. (B) The trajectory is extended into a discrete sequence of contact configurations using EFORT.}
  \label{fig:framework}
\end{figure*}

We claim that the desirable contact properties of a guide trajectory, proposed by \citeauthor{Bouyarmane2009}, can be formulated in a space of lower dimension, which we call $C_{reach}$. This formulation can make the planning of a guide trajectory more efficient computationally, while providing equivalent guarantees to planning directly in the configuration space. Among the particular properties obtained when planning in $C_{reach}$, we would like to guarantee that any reduced trajectory can actually lead to a feasible sequence of contacts, in which case we say that the reduced trajectory is truly feasible. It is possible in theory to guarantee that any reduced trajectory is truly feasible, even if it is more efficient in practice to approximate this property. The true-feasibility of the guide trajectory then allows us to focus on the selection of one particular sequence of contacts, for example one that minimizes the number of contacts in the sequence or maximizes the robot efficiency or style.

Based on these fundamental observations, we implement a very efficient acyclic contact planner. Our method is based on a probabilistic roadmap (PRM), that computes offline guide trajectories that are approximately truly feasible. The planner then resolves online the contact sequence by refining a guide trajectory computed from the PRM. Our planner is able to compute physically-consistent contact sequences for very complex systems (a humanoid, 28 joints; and an insectoid, 48 joints) in a few seconds for classical scenarios like climbing, and less than a minute for very complex problems like egress from a damaged truck. The planner also generalizes to planning dexterous manipulation movements, as demonstrated by preliminary results.

The contributions of the paper are twofold. We propose the first theoretical characterization of today's most efficient practical approach to sampled-based planning of acyclic contacts. And based on this characterization, we propose a very efficient and general implementation of an acyclic contact planner, the first one compatible with interactive applications.

We propose a framework to address the motion planning problem for multiped robots in cluttered environments: given a start and a goal configuration, the objective is to compute a sequence of contact configurations allowing to achieve the motion. For instance, we can consider the task of standing up, illustrated in Figure~\ref{fig:framework}--right. The problem is decoupled into two sequential phases: 1) the computation of a guide trajectory for the root of the robot; 2) the computation of a discrete sequence of contact configurations allowing to achieve the motion along the trajectory.
%
The remainder of this section presents the general organization of our method in Section~2. The two following sections 2 and 3 present respectively our answer to problems \Pa and \Pb. Finally, we propose a complete experimental validation of the planner with three very different kinematic chains (humanoid, insectoid and three-finger manipulator) in various scenarios.

>>>>

The issue of planning acyclic contacts was first completely described by \citeauthor{conf/iser/BretlRLKA04} in their seminal paper. The issue requires the simultaneous handling of two problems, $\mathcal{P}_1$: a relevant guide trajectory for the root of the robot in $SE(3)$; and $\mathcal{P}_2$: the planning of a discrete sequence of acyclic, balanced contact configurations along the trajectory (A third non trivial problem, $\mathcal{P}_3$, not adressed in this work, then consists in interpolating a complete motion between two postures of the contact sequence).  A key issue is to avoid combinatorial explosion when considering at the same time the possible contact configurations and the potential trajectories. This seminal paper proposes a first effective algorithm, able to handle simple situations (such as climbing scenarios), but not scalable to arbitrary environments. Following it, seve\-ral papers have applied this approach in particular situations, typically limiting the combinatorial by imposing a fixed set of possible contacts \citep{Hauser06usingmotion, stilman2010}.

Most of the papers that followed the work of \citeauthor{conf/iser/BretlRLKA04} have explored alternative formulations to handle the combinatorial issue. Two main directions have been explored. \textbf{On one hand, local optimization of both the root trajectory \Pa and the contact positions $\mathcal{P}_2$} has been used, to trade the combinatorial of the complete problem for a differential complexity, at the cost of local convergence. A complete example of the potential offered by such approaches was proposed \citep{Mordatch:2012:DCB:2185520.2185539} and successfully applied to a real robot \citep{mordatch2015}. To keep reasonable computation times, the method uses a simplified dynamic model for the avatar. Still, the computation time is far from interactive  (about 1 minute of computation for a sequence of 20 contacts). \citeauthor{DBLP:conf/humanoids/DeitsT14} propose to solve contact planning globally as a mixed integer problem, but only cyclic, bipedal locomotion is considered. Aside from the computation cost, a major drawback of these optimization based approaches is thus that they only offer local convergence when applied to acyclic contact planning.

@@

It thus appears that an integrated optimization based approach \citep{Mordatch:2012:DCB:2185520.2185539}, able to generate a locally optimal, complete trajectory within minutes, currently outperforms
existing planners, decoupled or not. However, recent contributions are able to interpolate dynamically executable 
trajectories between contact configurations \citep{herzog2015trajectory, Carpentier2016}. \citeauthor{Carpentier2016} are able to achieve this with real time performances.
To generate the input contact configurations, there is a need for an efficient contact planner, able to break the combinatorial to generate discrete contact sequences rapidly. 
Such a planner holds the promise of a complete real time multi contact locomotion system.

\subsection{Paper contribution and organization}
This paper presents a pragmatic approach to break the complexity of multi contact planning. With that objective in mind,
we believe that the separation between the guide trajectory and the contact sequence is the most promising direction \citep{DBLP:conf/iser/EscandeKMG08}.
However, this direction raises two theoretical questions that remain to be solved, or even to be properly formulated:
\begin{itemize}
\item The guide trajectory must satisfy a property guaranteeing the existence of a contact sequence to actuate it\endnote{This property is related to the controllability of the root actuated by the contact forces, but for discrete bounded properties.}. We call this property \textit{true feasibility}. This property has not been studied yet: the only way to validate a waypoint in the trajectory is to explicitly compute the contacts, which is computationally not reasonable \citep{Bouyarmane2009}.
\item There is an infinite combination of possible contact sequences for a given root trajectory. The selection of one particular contact sequence with interesting properties (minimum number of contact change, robustness, efficiency or naturalness) has been studied for cyclic cases \citep{Hauser06usingmotion}, but has not been efficiently applied to cluttered environments (\citeauthor{bouyarmane:lirmm-00777727, DBLP:conf/iser/EscandeKMG08} mostly randomly pick one contact sequence, leading to possibly very tedious contact sequences).
\end{itemize}

By tackling these issues, our method proposes contributions to both problems \Pa and \Pb.
For each problem, we introduce an original formulation of the related theoritical issues.
Then, we propose a concrete implementation of the solution, where the constraints of the problem are relaxed in favor of efficiency.
The relaxation is justified empirically, with statistics presented for each robot and scenario presented. The complete framework is implemented within
the open source platform Hpp.

To address \Pa we rigorously formulate a geometric condition for the true feasibility of a path, the reachability condition. We claim that this 
condition can be verified in a low dimensional space, that we call $C_{reach}$. This formulation can make the planning of a guide trajectory more efficient computationally, while providing equivalent guarantees to planning directly in the configuration space.
In practice, for each considered robot, we compute an approximation of $C_{reach}$. From this approximation,
an efficient trade-off between a necessary and a sufficient condition for true feasibility is obtained. This trade-off allows the implementation of a low-dimensional, Reachability-Based RRT planner (RB-RRT), able to find truly feasible paths for the root of the robot within a few milliseconds to a few seconds depending on the scenario.

To address \Pb, we consider a truly feasible path as an input, and generate new contact configurations along it using a sampling approach.
The true feasibility of the path provides guarantees that the contacts can be created, and reduce the dimensionality of the problem (at each step the root position is fixed). The theorical completeness of the approach is partially traded for efficiency: instead of sampling online
limb configurations and projecting them onto contact, we precompute offline a large database of limb configurations, stored in a spatial structure optimized
for proximity queries. At runtime, this framework allows to obtain extremly rapidly a set of candidate configurations for contact.
These candidates are sorted using a criterion to evaluate the quasi static balance of the configuration, and ranked according to user defined heuristics.
In our implementation we use two such heuristics, used in all our experiments: a robustness criterion regarding the balance of a configuration, and a manipulability based heuristic,
the Extended FORce Transmission ratio (EFORT).

The contributions of the paper are thus twofold. We propose a theoretical characterization of today's most efficient practical approach to sampled-based planning of acyclic contacts. Based on this characterization, we propose a very efficient and general implementation of an acyclic contact planner, the first one compatible with interactive applications.

In comparison with previous work, our planner does not produce a complete motion, but a discrete sequence of contacts.
Because we showed that the interpolation of a complete trajectory between such configurations can be solved within real time constraints \citep{Carpentier2016}, we claim that our planner
is the fastest multi-contact planning solution to our knowlege.

The present paper is an extension of a conference paper to appear in the proceedings of the ISRR '15 conference TODO CITE.
The conference paper focuses on the theoritical formulation of the problem, and presents results obtained with virtual avatars.
This extension completes this work with an argumented discussion on the implementation of this theory to real world robots and problems.
In particular, we detail the robust balance criteria designed to ensure the equilibrium of the robot. Using HRP-2 as a running example, we describe how $C_{reach}$ is approximated to address
\Pa, and provide the pseudo code of the contact generation algorithm implemented to find balanced contact sequences.

In Section~\ref{overview}, we present the general organization of our method. Section~\ref{rbprm} and Section~\ref{sec:efort} present respectively our answer to problems \Pa and \Pb. In Section~\ref{heuristics}, we present two heuristics used for the selection of the best contact configuration. Finally, we propose a complete experimental validation of the planner with three very different kinematic chains (the HRP-2 and Hyq robots, and a three-finger manipulator) in various scenarios,
that complete those presented in the conference paper.
